k-mer-based GWAS enhances the discovery of causal variants and candidate genes in soybean

### 2023-03-31 additional_files/manuscript.tex
Correct some typos upon re-reading the manuscript

% Packages added by M-A Lemay
\usepackage{booktabs} % prettier tables
\usepackage{threeparttable} % easy addition of footnotes below tables

@@

Given the current performance of SV genotyping programs, it is not clear
whether using SVs for GWAS provides any advantage over using \emph{k}-mer- or
SNP-based GWAS, or a combination of both.  Indeed, in the single case where the
SV-based analyses identified the causal variant, the \emph{k}-mer-based GWAS
identified it as well. Moreover, SNP-based GWAS provided better guidance
in delimiting the signals found at a given locus because of the much higher
density of markers compared to SVs.  The usefulness of SV-based GWAS compared
to SNP-based GWAS will largely depend on whether SVs can represent variation
that is not otherwise in LD with SNPs. In tomato, Domínguez et al. cite{dominguez2020} found
that most transposable element insertions associated with phenotypic variation
in agronomic traits were not tagged by SNPs. On the other hand, in humans,
Maretty et al. \cite{maretty2017} found that the majority of SVs were in high LD with SNPs. It
is therefore likely that how well SVs can be tagged by SNPs will depend on the
particular type of SV and on the species. A more efficient approach than
SV-based GWAS may be to conduct \emph{k}-mer-based GWAS and only then link
known SVs to significant \emph{k}-mers.  Alternatively, \textit{k}-mer-based
GWAS can be conducted specifically for \textit{k}-mers that are associated with
SVs \citep{jayakodi2020}.  Previous studies have obtained noteworthy results by
using SVs for GWAS analyses \citep[e.g.][]{zhang2015, akakpo2020,
dominguez2020}, however it would be worthwhile to see whether these results
could be replicated using \emph{k}-mer-based GWAS.

@@

To help with the interpretation of results and the identification of candidate
genes, we defined regions of highly significant associations within each signal
by delimiting a range of coordinates containing the top 5\% most significant
associations (Paragraph) or 1\% most significant associations (SNPs/indels
and \emph{k}-mers) found within the signal. In cases where such a fraction
represented fewer than two markers or \emph{k}-mers, the highly significant
region was simply delimited by the two most significant associations.

@@

\section*{Abbreviations}%% if any
BAM: Binary alignment map format;
CDW: corrected dry weight;
CHS: chalcone synthas;
CNV: copy number variant, copy number variation;
GRIN: Germplasm Resources Information Network;
GWAS: Genome-wide association study;
HPS: Hydrophobic protein from soybean;
LD: Linkage disequilibrium;
MAF: Minor allele frequency;
MATE: ultidrug and toxic compound extrusion;
MLM: Mixed linear model;
NCBI: National Center for Biotechnology Information;
SNP: Single-nucleotide polymorphism;
SRA: Sequence Read Archive;
SV: Structural variant, structural variation;
VCF: Variant call format;
WGS: Whole-genome (re)sequencing

>>>>

% Packages added by M-A Lemay
\usepackage{booktabs} % prettier tables
\usepackage{threeparttable} % easy addition of footnotes below tables
\usepackage[T1]{fontenc}

@@

Given the current performance of SV genotyping programs, it is not clear
whether using SVs for GWAS provides any advantage over using \emph{k}-mer- or
SNP-based GWAS, or a combination of both.  Indeed, in the single case where the
SV-based analyses identified the causal variant, the \emph{k}-mer-based GWAS
identified it as well. Moreover, SNP-based GWAS provided better guidance
in delimiting the signals found at a given locus because of the much higher
density of markers compared to SVs.  The usefulness of SV-based GWAS compared
to SNP-based GWAS will largely depend on whether SVs can represent variation
that is not otherwise in LD with SNPs. In tomato, Domínguez et al. \cite{dominguez2020} found
that most transposable element insertions associated with phenotypic variation
in agronomic traits were not tagged by SNPs. On the other hand, in humans,
Maretty et al. \cite{maretty2017} found that the majority of SVs were in high LD with SNPs. It
is therefore likely that how well SVs can be tagged by SNPs will depend on the
particular type of SV and on the species. A more efficient approach than
SV-based GWAS may be to conduct \emph{k}-mer-based GWAS and only then link
known SVs to significant \emph{k}-mers.  Alternatively, \textit{k}-mer-based
GWAS can be conducted specifically for \textit{k}-mers that are associated with
SVs \citep{jayakodi2020}.  Previous studies have obtained noteworthy results by
using SVs for GWAS analyses \citep[e.g.][]{zhang2015, akakpo2020,
dominguez2020}, however it would be worthwhile to see whether these results
could be replicated using \emph{k}-mer-based GWAS.

@@

To help with the interpretation of results and the identification of candidate
genes, we defined regions of highly significant associations within each signal
by delimiting a range of coordinates containing the top 5\% most significant
associations (SVs) or 1\% most significant associations (SNPs/indels
and \emph{k}-mers) found within the signal. In cases where such a fraction
represented fewer than two markers or \emph{k}-mers, the highly significant
region was simply delimited by the two most significant associations.

@@

\section*{Abbreviations}%% if any
BAM: Binary alignment map format;
CDW: corrected dry weight;
CHS: chalcone synthase;
CNV: copy number variant, copy number variation;
GRIN: Germplasm Resources Information Network;
GWAS: Genome-wide association study;
HPS: Hydrophobic protein from soybean;
LD: Linkage disequilibrium;
MAF: Minor allele frequency;
MATE: Multidrug and toxic compound extrusion;
MLM: Mixed linear model;
NCBI: National Center for Biotechnology Information;
SNP: Single-nucleotide polymorphism;
SRA: Sequence Read Archive;
SV: Structural variant, structural variation;
VCF: Variant call format;
WGS: Whole-genome (re)sequencing
